---
title: "Problem Set 1"
author: "Seungho Choi"
date: "2025-09-13"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r}
library(tibble)
library(tidyverse)
```


## 3. Starting in R

a) generate some simulated data

```{r}
X = rnorm(n = 1000, mean = 0, sd = 1)
X[1]
```


b) Calculate the following sample moments of the data you have generated:
  
```{r}
mean(X)
sd(X)
sd(X)
```


c) Construct the running mean of the random variables:
  
```{r}
running_mean <- cumsum(X) / seq_along(X)
library(purrr)
library(ggplot2)
data <- tibble(running_mean) %>%
  rowid_to_column()
ggplot(data, aes(x=rowid, y=running_mean)) +
  geom_point() + 
  geom_line(aes(y=0))
```


d) Repeat the exercise for σ = 2 and σ = .2

```{r}
X = rnorm(n = 1000, mean = 0, sd = 2)
X[1]

mean(X)
sd(X)
sd(X)

running_mean <- cumsum(X) / seq_along(X)
library(purrr)
library(ggplot2)
data <- tibble(running_mean) %>%
  rowid_to_column()
ggplot(data, aes(x=rowid, y=running_mean)) +
  geom_point() + 
  geom_line(aes(y=0))
```
```{r}
X = rnorm(n = 1000, mean = 0, sd = 0.2)
X[1]

mean(X)
sd(X)
sd(X)

running_mean <- cumsum(X) / seq_along(X)
library(purrr)
library(ggplot2)
data <- tibble(running_mean) %>%
  rowid_to_column()
ggplot(data, aes(x=rowid, y=running_mean)) +
  geom_point() + 
  geom_line(aes(y=0))
```


Does the sample mean seem to converge faster, slower, or independently of the variance?
  
The sample mean seems to converge faster when it has a lower variance.

To explain this in more detail, the consequence of plotting different values for the standard deviation is clearly shown by the graph. When I changed the value to 2, instead of 1, not only did the range of the y-axis (running_mean) expand, but also, unlike the original graph, where the fluctuation of the line becomes steadied as it gets closer to the end, this new graph showed a nonstop fluctuation. On the other hand, in the case where I put 0.2 instead of 1, although the initial starting point is still fluctuating, shortly after the first part, I was able to see that the variance of the entire graph is a lot steadier than the two graphs that I have created previously.